LaMachine: A meta-distribution for NLP software

### 2019-03-29 docs/lamachine_clarin2018.tex
added  to clarin paper

LaMachine exists since May 2015 and has been installed extensively ever since by numerous users. Quantifying its actual
use is not trivial despite having sme tracking solutions in place, but our statistics show that LaMachine was installed at least over a thousand times, on
machines in over 25 different countries since its inception\footnote{https://applejack.science.ru.nl/lamastats/lamachinestats.html}. In early 2018 version 2 was released which was a significant redesign, powered
by Ansible\footnote{\url{https://www.ansible.com}}. We ourselves use LaMachine as a means by to make our full software
stack available to end-users, whether those be individual researchers, research groups, or hosting providers like CLARIN
centres. LaMachine has in fact been adopted in all of those areas. We also see a regular influx of questions on our issue tracker, indicating community interest.

@@

@@

@@

LaMachine is most often used by a researcher or data scientist who installs LaMachine on his own computer (locally, in a
local environment or as a Docker container or VM). Here we discuss two case studies where LaMachine was used under
circumstances that go beyond this most common usage of LaMachine: using LaMachine on a secure server without internet
connection, and using LaMachine to provide a work environment for a group of researchers.

@@

We used a virtual machine (VM) image of LaMachine and we ran this 64-bits Linux-based VM inside another
VM with Windows Server 2012, provided to us by the health inspectorate for this project, in which we did have administrative
rights but no internet access. In terms of hardware we ran on a machine with 8 cores and  32GB internal memory available.
%using VirtualBox and Vagrnt on the Windows VM
LaMachine provided a fully functional research environment and we ran all our experiments within LaMachine. We interacted with LaMachine both through the command line, which offers a standard
shell and enables access to all lower-level tools and programming languages; and through the (offline) webbrowser to use
the Jupyter Notebook environment.

@@

LaMachine offered a convenient platform for a range of different explorations and experiments in the area of NLP and
text mining. However, for some situations LaMachine, or rather Linux in general, was not a good fit for the audience of
the workshop: For team members who did not have experience with a non-Windows environment, LaMachine was not a suitable
or useful tool. The limit of LaMachine was also reached for members who wanted to use desktop text editors with a
graphical user interface as this is not offered by LaMachine. Moreover, we did not manage to get X-forwarding working in the
Ubuntu Linux VM and after a few attempts the team gave up on resolving this issue due to time pressure. This, also
demonstrates that fine-tuning the configuration of certain aspects of LaMachine, but especially beyond LaMachine, is beyond the reach of a data scientist
without system administration skills. This certainly also applies also to the installation as a whole in the
SURFsara context, which involved things like the partitioning, formatting and mounting of (virtual) drives and setting
up user accounts on the shared VM, all of which require some system administration skills and are too context-specific
to be within the scope of LaMachine.
LaMachine was convenient and speeded up writing code as the most common scientific data-related packages are already present in LaMachine.
 %verwerkt <-- <iris> it moet nog wat beter geformuleerd worden maar je snaprt wat ik bedoel
%@iris: hebben jullie verder geen gebruik proberen te maken van de Jupyter Lab omgeving? Dat biedt in ieder geval iets
%grafisch wat via web te benaderen is?
%@maarten helaas lukte dat ook niet, dat liep bij mij steeds vast. - ik heb opmerking van alex over tijdswinst toegevoefd

@@

Use cases as the examples in section \ref{sec:case} contribute to thorough testing and running of LaMachine in less ideal
circumstances such as nested VM constructions and offline usage.

>>>>

LaMachine exists since May 2015 and has been installed extensively ever since by numerous users. Quantifying its actual
use is not trivial despite having some tracking solutions in place, but our statistics show that LaMachine was installed at least over a thousand times, on
machines in over 25 different countries since its inception\footnote{https://applejack.science.ru.nl/lamastats/lamachinestats.html}. In early 2018 version 2 was released which was a significant redesign, powered
by Ansible\footnote{\url{https://www.ansible.com}}. We ourselves use LaMachine as a means by to make our full software
stack available to end-users, whether those be individual researchers, research groups, or hosting providers like CLARIN
centres. LaMachine has in fact been adopted in all of those areas. We also see a regular influx of questions on our issue tracker, indicating community interest.

@@

The CLARIN Switchboard \cite{zinn2016clarin} also offers access to NLP tools at a different level as it provides a webinterface where users can upload their texts for NLP processing, proving very easy acces. Lamachine is most often used by data scientists and reseachers who do have programming experience and want to work with complex NLP software.

@@

Virtual Research Environments (VRE) such as the Dutch CLARIAH WP3 VRE project\footnote{\url{https://github.com/meertensinstituut/clariah-wp3-vre}} and the Spanish VRE project \cite{gonzalez2017evi} aim to develop more high level tool environments with graphical interfaces to accomodate the reseacher. Lamachine is currently developed in close collaboration with the Dutch CLARIAH WP3 VRE project.

@@

LaMachine is most often used by a researcher or data scientist who installs LaMachine on his own computer (locally, in a
local environment or as a Docker container or VM). Here we discuss two case studies where LaMachine was used under
circumstances that go beyond this most common usage of LaMachine: using LaMachine on a secure server without internet
connection, and using LaMachine to provide a work environment for a group of researchers. These case studies illustrate the advantages and limitations of Lamachine in practice.

@@

We used a virtual machine (VM) image of LaMachine and we ran this 64-bits Linux-based VM inside another
VM with Windows Server 2012, provided to us by the health inspectorate for this project, in which we did have administrative
rights but no internet access. In terms of hardware we ran on a machine with 8 cores and  32GB internal memory available. Note that as we had no internet access available on the working server, we prepared a custom-made VM image of Lamachine that included an additional editor and some extra Python libraries.
%using VirtualBox and Vagrnt on the Windows VM
LaMachine provided a fully functional research environment and we ran all our experiments within LaMachine. We interacted with LaMachine both through the command line, which offers a standard
shell and enables access to all lower-level tools and programming languages; and through the (offline) webbrowser to use
the Jupyter Notebook environment.

@@

LaMachine offered a convenient platform for a range of different explorations and experiments in the area of NLP and
text mining. However, for some situations LaMachine, or rather Linux in general, was not a good fit for the audience of
the workshop: for team members who did not have experience with a non-Windows environment, LaMachine was not a suitable
or useful tool. The limit of LaMachine was also reached for members who wanted to use desktop text editors with a
graphical user interface as this is not offered by LaMachine. Moreover, we did not manage to get X-forwarding working in the
Ubuntu Linux VM and after a few attempts the team gave up on resolving this issue due to time pressure. This, also
demonstrates that fine-tuning the configuration of certain aspects of LaMachine, but especially beyond LaMachine, is beyond the reach of a data scientist
without system administration skills. This certainly also applies also to the installation as a whole in the
SURFsara context, which involved things like the partitioning, formatting and mounting of (virtual) drives and setting
up user accounts on the shared VM, all of which require some system administration skills and are too context-specific
to be within the scope of LaMachine.
LaMachine was convenient and speeded up writing code as the most common scientific data-related packages are already present in LaMachine.
 %verwerkt <-- <iris> it moet nog wat beter geformuleerd worden maar je snaprt wat ik bedoel
%@iris: hebben jullie verder geen gebruik proberen te maken van de Jupyter Lab omgeving? Dat biedt in ieder geval iets
%grafisch wat via web te benaderen is?
%@maarten helaas lukte dat ook niet, dat liep bij mij steeds vast. - ik heb opmerking van alex over tijdswinst toegevoefd

@@

Most users of Lamachine are installing the software distribution in their own machine, either locally, via Docker or VM. However, we described to use cases in section \ref{sec:case} where we explore the possibilities under more constrained and demanding circumstances.
Such case studies contribute to thorough testing and running of LaMachine in less ideal circumstances such as nested VM constructions and offline usage.
